---
title: "Lab3_JunLi_0"
subtitle: "Bioinformatics -- 732A51"
author: "Jun Li"
date: '2020-12-07'
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval=TRUE,echo=F,
                      message = F, warning = F, error = F,
                      fig.width=7, fig.height=5, fig.align="center")
```

```{r}
library(ape)
library(phyloch)

library(ade4)
library(mvMORPH)
library(mvSLOUCH)
library(ouch)
library(slouch)
library(rms)
```


## Question 1
### part 1.1


```{r,echo=T}
### part 1.1
##in order to install phyloch
## Sys.setenv(R_REMOTES_NO_ERRORS_FROM_WARNINGS=TRUE)
## devtools::install_github("fmichonneau/phyloch",dependencies=T)

x <- paste("AJ5345", 26:49, sep = "")
x <- c("Z73494", x)
sylvia.seq <- read.GenBank(x)  ## download DNA sequences
sylvia.clus <- clustal(sylvia.seq)  ## align DNA sequences

taxa.sylvia <- attr(sylvia.seq, "species")
names(taxa.sylvia) <- names(sylvia.seq)
rm(sylvia.seq)
taxa.sylvia[1] <- "Sylvia_atricapilla"
taxa.sylvia[24] <- "Sylvia_abyssinica"

#syl.K80 <- dist.dna(sylvia.clus, pairwise.deletion = TRUE) ##matrix of pairwise distances, using model K80 
#nj.sylvia.K80 <- nj(syl.K80)  ## fit a phylogenetic tree based on distances

f <- function(xx) root(nj(dist.dna(xx, p=TRUE)), "AJ534526") ## tree estimation function
tr <- f(sylvia.clus) ## fit a tree
nj.boot.sylvia <- boot.phylo(tr, sylvia.clus, f, 200,rooted = TRUE) ## bootstrap analysis

tr$tip.label <- taxa.sylvia[tr$tip.label] ## rename tree tips

plot(tr, no.margin = TRUE) 
nodelabels(round(nj.boot.sylvia / 200, 2), bg = "white") ## label tree nodes
add.scale.bar(length = 0.01)
#write.tree(nj.est, "sylvia_nj_k80.tre")


tr <- drop.tip(tr, "Chamaea_fasciata") ## drop missing-data tip
sylvia.eco <- read.table("sylvia_data.txt")
DF <- sylvia.eco[tr$tip.label, ] ## reorder geography data 

```



### part 1.2
It shows that the SYM model gives the highest likelihood, therefore considered as the optimal model in this case. And the most likely ancester comes from "temptrop" region. Transition between trop and temptrop has the highest rate of 14.8 with higher variance, temptrop and temp of 4.9 with lower variance, while trop and temp of almost zero with unknown variance. Conclusion is that the ancestor dates back from temptrop region, then larger proportion migrates to trop and less to temp, and no interleves between trop and temp.



```{r}
### part 1.2
syl.er <- ace(DF$geo.range, tr, type="d", model="ER")
syl.ard <- ace(DF$geo.range, tr, type="d", model="ARD")
syl.sym <- ace(DF$geo.range, tr, type="d", model="SYM")

print("Here comes the estimation by ER:")
syl.er
print("Here comes the estimation by ARD:")
syl.ard
print("Here comes the estimation by SYM:")
syl.sym
```


## Question 2
### part 2.1
A phylogeny tree, histograms and scatter plot for size and range data is shown as below. Species have mean size of 14.3, max size of 266.5 and min size of 0.04.


```{r}
### part 2.1
data(carni70)
tre<-carni70$tre
tab<-carni70$tab

tree<-as.phylo(newick2phylog(tre))
plot(tree)  ## phylogeny tree

hist(tab$size,breaks=100)
hist(tab$range,breaks=100)
plot(tab$size,tab$range)
#tree_root<-root(tree,"Puma_concolor")
#ace.sym <- ace(tab$range, tree, type="c")
#print("Here comes the estimation by SYM:")
#ace.sym
```

### part 2.2
Results from different models are as below:

Model 1: AIC value for fitting variable size and range is 652 respective 533.

Model 2: AIC value 1187

Model 3: AIC value for fitting variable size and range is 654 respective 503

Model 4: AIC value 1161

Model 5: AIC value 1205

Therefore, Model 3 fitting traits as independent OU processes has the lowest AIC value, and should be the optimal one.

```{r}
### part 2.2

x<-tab$size
y<-tab$range
tr1<-multi2di(tree)
tab=tab[tr1$tip.label,]

print("Model 1:")
#pic.x<-pic(x,phy = tr1)
#pic.y<-pic(y,phy = tr1)
#cor.test(pic.x,pic.y)
#p-value is 0.3481, therefore hypothesis "true correlation is equal to 0" can not be rejected, thus two traits are independent, which conforms with the assumed model
#summary(lm(pic.y~pic.x-1))

fit11<-mvBM(tr1, tab$size, model="BM1", method="pic")
fit12<-mvBM(tr1, tab$range, model="BM1", method="pic")
#AIC(fit11);AIC(fit12)
#LRT(fit11,fit12)
``` 





```{r}
print("Model 2:")
#bm<- corBrownian(phy=tr1)
#fit2<-Gls(range ~ size, correlation=bm, data=as.data.frame(tab))
#fit2
#print(paste("AIC value is:",AIC(fit2)))
#P-value is 0.409 for intercept and 0.35 for size coefficient. Therefore there is no significant linear relationship between traits, which conflicts with the assumed model.

fit2<-mvBM(tr1, tab, model="BM1", method="pic")
```





```{r}
print("Model 3:")

fit31<- mvOU(tree=tr1,data=tab$size,  model="OU1")
fit32<- mvOU(tree=tr1,data=tab$range,  model="OU1")
#LRT(fit31,fit32)
#AIC(fit31);AIC(fit32)

```

Model 4:



```{r}
print("Model 4:")
fit4<- mvOU(tree=tr1,data=tab,  model="OU1")

```

Model 5:

```{r}
print("Model 5:")
fit5<-mvslouchModel(tr1,mData=as.matrix(tab),kY=1)
#fit51<-mvBM(tr1, tab$size, model="BM1", method="pic")
#fit52<-mvOU(tree=tr1,data=tab$range,  model="OU1")
#LRT(fit51,fit52)
fit5
```






## Appendix code
```{r ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```